TGI-spinal-power

### 2022-05-17 analysis/tgispine_zoib_power-sim.Rmd
creating model and simulation functions

# Simulate data for different N (n = 20; n = 30; n = 40)
```{r}
# creating individual data-frames from simulated data
n = 20
df20 <- data.frame()
for (i in 1:n) {
  df <- simulate_subj(i, 0, 1)
  df20 <- rbind(df20,df)}

n = 30
df30 <- data.frame()
for (i in 1:n) {
  df <- simulate_subj(i, 0, 1)
  df30 <- rbind(df30,df)}

n = 40
df40 <- data.frame()
for (i in 1:n) {
  df <- simulate_subj(i, 0, 1)
  df40 <- rbind(df40,df)}
df20 #just show a sample, to check
```

# Plotting data distributions to see what kind of model family is required
```{r}
# do this for the largest simulated dataset
df40$quality <- factor(df40$quality, 
                             levels = c("cold", "warm", "burn"))
ggplot2::ggplot(data = df40) +
  geom_density(aes(rating, fill = stimuli), alpha = .5) +
  facet_wrap(~quality) +
  theme_classic()
```
# Summary statistics for the largest data set - for plotting
```{r}
# define each condition
df40$conditionN[df40$dermatome == "within" & df40$cold_location == "dist_rostr"] <- 1
df40$conditionN[df40$dermatome == "within" & df40$cold_location == "prox_caud"] <- 2
df40$conditionN[df40$dermatome == "across" & df40$cold_location == "dist_rostr"] <- 3
df40$conditionN[df40$dermatome == "across" & df40$cold_location == "prox_caud"] <- 4

@@

# Now to model effects! Try a zero inflated regression model
```{r}
# First model cold VAS
model.cold40 = glmmTMB::glmmTMB(rating ~ stimuli + dermatome + cold_location + trialn +
                                (1|id),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+stimuli,
                      data = filter(df40, quality == 'cold'),
                      na.action = na.omit) 
# then model warm VAS
model.warm40 = glmmTMB::glmmTMB(rating ~ stimuli + dermatome + cold_location + trialn +
                                (1|id),
                      family = glmmTMB::beta_family(),
                      ziformula = ~1+stimuli,
                      data = filter(df40, quality == 'warm'),
                      na.action = na.omit) 
# model summaries
summary(model.cold40)
summary(model.warm40)
```
# Look at model assumptions
```{r}
# cold assumption - on n = 20
model.cold.assmup <- simulateResiduals(model.cold40, n = 1000)
plot(model.cold.assmup)
```

```{r}
# warm assumption - on n = 20
model.warm.assump <- simulateResiduals(model.warm40, n = 1000)
plot(model.warm.assump)
```

# Extract the power from each model
```{r}

>>>>

# Plotting data distributions to see what kind of model family is required
```{r}
# do this for the largest simulated dataset
df40$quality <- factor(df40$quality, 
                             levels = c("cold", "warm", "burn"))
ggplot2::ggplot(data = df40) +
  geom_density(aes(rating, fill = stimuli), alpha = .5) +
  facet_wrap(~quality) +
  theme_classic()
```
# Model function - Using zero inflated regression model
```{r}
# First model cold VAS
model <- function(x, vas = c('cold','warm','burn')) {
  if (vas[1] == 'cold'){
    x1 <- x %>%  filter(quality == 'cold')
  } else if (vas[1] == 'warm'){

  } else if (vas[1] == 'burn') {
    x1 <- x %>%  filter(quality == 'burn')
  }

  mod = glmmTMB::glmmTMB(
    rating ~ stimuli + dermatome + cold_location + trialn + (1|id),
    family = glmmTMB::beta_family(),
    ziformula = ~1+stimuli,
    data = x1,
    na.action = na.omit)

return(summary(mod))}
```

# Extract the power from each model
```{r}
# the warm model has the smallest effect size, so should extract power calculation for this model only
# do this for 40 participants, for 1000 simulations
nSims <- 1000
po <- data.frame()
for (i in 1:nSims){
  # creating individual data-frames from simulated data
  n = 40
  df40 <- data.frame()
  for (i in 1:n) {
    df <- simulate_subj(i, 0, 1)
    df40 <- rbind(df40,df)}
  # then run the model based on simulated data
  m <- model(df40, 'warm')
  # extracting the p-value and effect size
  tmp <- c(m$coefficients$cond[4,4], m$coefficients$cond[4,1])
  po <- rbind(po,tmp)
}

# rename columns
names(po)[1] <- 'pval'
names(po)[2] <- 'eff'
po$sig <- as.numeric(po$pval < .05) #is significant
po$sig[is.na(po$sig)] <- 0 #removing any NAs (code as unsig)

power <- sum(po$sig)/nSims
power
```

# Now plot most recently simulated data
# Summary statistics for the largest data set - for plotting
```{r}
# define each condition
df40$conditionN[df40$dermatome == "within" & df40$cold_location == "dist_rostr"] <- 1
df40$conditionN[df40$dermatome == "within" & df40$cold_location == "prox_caud"] <- 2
df40$conditionN[df40$dermatome == "across" & df40$cold_location == "dist_rostr"] <- 3
df40$conditionN[df40$dermatome == "across" & df40$cold_location == "prox_caud"] <- 4

@@

# Look at model assumptions
```{r}
# cold assumption - on n = 20
model.cold.assmup <- simulateResiduals(model.cold40, n = 1000)
plot(model.cold.assmup)
```
